Combined Data Cleaning and Exploratory Analysis Script for Manuscript: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health

### 2023-04-26 3. Analysis Script/StatisticalAnalysis.qmd
Edits Continued (Analysis)

## **Loading Packages**
```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels)
library(tidyr)
library(gtsummary)
library(dplyr)
library(vip)
library(performance)
library(jtools)
library(broom)
library(broom.mixed)
```

@@

lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Table, PT_Data)

@@

@@

lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)

@@

@@

plot_summs(lmall_PT, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
```

@@

@@

plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
```
# **IP Analysis KP finish by 4/9!!**

#**Time to event**
```{r}
#KP needs to edit this! Not running as expected....

library(finalfit)
library(forcats)
library(survival)

@@

#**Cox regression: Picnic Table Data**
```{r}

@@

#Date needs to be fixed in this DS

FD_Data <- FD_Data %>%
  mutate(Collection_Date=lubridate::dmy(Collection_Date))

@@

coxph(Surv(Collection_Date, Absence_0_Presence_1 ~ Sample_Type, FD_Data))

coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
  tbl_regression(exp=TRUE)
```

>>>>

## **Loading Packages**
```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(dotwhisker)
library(here) #for data loading/saving
library(tidymodels)
library(tidyr)
library(gtsummary)
library(dplyr)
library(vip)
library(performance)
library(jtools)
library(broom)
library(broom.mixed)
```

@@

lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Date, PT_Data)

@@

#Checking model performance: Picnic Table
check_model(lmfit_PT$fit)

tidy(lmfit_PT) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))

@@

model_lm<- linear_reg()

lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Sample_Type+Date, PT_Data)

@@

#Checking model performance: Picnic Table
check_model(lmfit_PT1$fit)

tidy(lmfit_PT1) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))

@@

#Do not want all predictors...

#Checking model performance: Picnic Table
check_model(lmall_PT$fit)
```

@@

#Checking model performance: FEEDER
check_model(lmfit_FD$fit)

@@

plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
```
# **Generalized Linear Model: Picnic Table** 
```{r}
lmog<-linear_reg() %>% 
  set_engine("glm")

lmfit_PT_glm<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Table, PT_Data)

lmfit_PT_glm

glimpse(lmfit_PT_glm)

#Way 1
glance(lmfit_PT_glm)

#Way 2
tidy(lmfit_PT_glm)

plot_summs(lmfit_PT_glm)

#Fitting other specific indicators
lmog_glm1<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)

lmfit_PTglm1

glimpse(lmfit_PTglm1)

#Way 1
glance(lmfit_PTglm1)

#Way 2
tidy(lmfit_PTglm1)

plot_summs(lmfit_PTglm1)

#Fitting other specific indicators
lmog_glm2<-linear_reg() %>% 
  set_engine("glm")

lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Quadrant+Table, PT_Data)

lmfit_PTglm2

glimpse(lmfit_PTglm2)

#Way 1
glance(lmfit_PTglm2)

#Way 2
tidy(lmfit_PTglm2)

plot_summs(lmfit_PTglm2)

#Trying ALL predictors of Picnic Table dataset 
lmog_PTGLM<- linear_reg() %>%
  set_engine("glm")

lmall_PTGLM<- lmog_PTGLM %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PTGLM

tidy(lmall_PTGLM)
```
# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
```{r}
PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)

PT_GLM_PER

PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial)

PT_GLM_PREV
```
# **Generalized Linear Model: Feeder Data**
```{r}
lmog_FD<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_glm<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD_glm

glimpse(lmfit_FD_glm)

#Way 1
glance(lmfit_FD_glm)

#Way 2
tidy(lmfit_FD_glm)

plot_summs(lmfit_FD_glm)

#Fitting other specific indicators
lmog_FD_GLM<-linear_reg() %>% 
  set_engine("glm")

lmfit_FD_GLM<- lmog_FD_GLM %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD_GLM

glimpse(lmfit_FD_GLM)

#Way 1
glance(lmfit_FD_GLM)

#Way 2
tidy(lmfit_FD_GLM)

plot_summs(lmfit_FD_GLM)

#Trying ALL predictors of Picnic Table dataset 
lmog_FDGLM_all<- linear_reg() %>%
  set_engine("glm")

lmall_FDGLM_all1<- lmog_FDGLM_all %>% #Training 
  fit(Absence_0_Presence_1~., FD_Data)

lmall_FDGLM_all1

tidy(lmall_FDGLM_all1)
```

# **Generalized Linear Model: Feeder Data (Persistence & Prevalence)**
```{r}
FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial)

FD_GLM_PER

FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial)

FD_GLM_PREV
```
#**Time to event**
```{r}
library(finalfit)
library(forcats)
library(survival)

@@

#**Cox regression: Picnic Table Data**
```{r}
PT_Data$Date<- as.numeric(PT_Data$Date)

@@

@@

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data)

coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
  tbl_regression(exp=TRUE)
```

### 2023-04-26 3. Analysis Script/exploratory_analysis.qmd
Edits Continued (Analysis)

PTD_Quad<- ggplot(PTD_clean, 
              aes(fill=Quadrant, y= Salmonella_Positive, x= Date)) + 
  geom_bar (position="stack", stat="identity") +
  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1) +
  ggtitle("Positive Salmonella Samples by Date and Quadrant") +
  labs(y= "Salmonella Positive", x= "Collection Date")

@@

FD_plot_final<- FD_plot + labs(fill="Feeder Type")

>>>>

PTD_Quad<- ggplot(PTD_clean, 
              aes(fill=Quadrant, y= Salmonella_Positive, x= Date)) + 
  geom_bar (position="stack", stat="identity") +
  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1) +
  ggtitle("Positive Salmonella Samples by Date and Quadrant") +
  labs(y= "Salmonella Positive", x= "Collection Date") +

@@

FD_plot_final<- FD_plot + labs(fill="Feeder Type") + scale_fill_discrete(labels=c('Plastic', 'Plastic Coated', 'Wood', 'Wood Coated'))
